Semantic Prediction in {R}ussian

### 2020-04-11 paper.tex
rework corpora and models and metrics language

We used the dataset with cloze task answers from \cite{laurinavichyute}. The dataset is based on 144 sentences randomly selected from the Russian National Corpus (RNC, ruscorpora.ru) - an online corpus of Russian texts with extensive search options. These sentences were slightly edited: the authors replaced rare infrequent words with more frequent ones and shortened the sentences when they exceeded the preset maximum length of 13 words. The stimuli sentences were subjected to the cloze task experiment. Respondents were asked to successively predict the next words for each context. An example of stimuli that were shown to the participants is presented in Table 1 with corresponding correct next words and calculated predictability scores.

@@

Russian news corpus \cite{runews} includes newspaper articles published in the 2000s: newspaper different types of articles.
National corpus of Russian language \cite{natcorp} includes written texts from the middle of the 18th to the middle of the 20th century.

Corpus characteristics are presented in Table 2.

@@

All of our models (except LSTM) were trained separately on 2 corpora and their combination. Thus we examine these models:

@@

Custom LSTM model was trained on the part of RNC, excluding the sentences chosen for stimuli, namely on modern written texts from the 1950s to the present day, real-life Russian oral speech recordings from the same period. Besides, it was trained on RNC and NCRL. Overall, the training corpus consisted of 577 million tokens. The model was tested on 1000 sentences from the OpenCorpora \cite{bocharov:2011} with an overall 1,9 million tokens from newspaper articles, Russian Wikipedia, texts from blogs, fiction, non-fiction, and legal documents.

@@

1)        Extracts first 10 words (we have decided that that is the maximum amount of words for each context that would have meaningful probabilities as all of the words after the first ten for each model have their probabilities tends to zero) for each of 1219 contexts;

% TODO arithmetic should be TeX-ified
2)        Counts the quotient of the sum of words’ probabilities and each amount of words (for example, if two words were to have probabilities 0.5 and 0.1, we would count $((0.5+0.1)/2)$, which we call mean probabilities);

3)        Counts the difference between the probabilities of the first and last word and then the difference between mean probabilities of the previous word and the next one;

4)        Decides what amount of words vectors to use for each context based on how different the first-last word difference and the mean probability difference is – if the latter is lesser than the former, the function checks the next difference, if not – the previous amount of words would be used for semantic comparison.

After the decision on the number of words was made, with the help of gensim in Python a vector for each word was extracted.

>>>>

We used the dataset with cloze task answers from \cite{laurinavichyute}. The dataset is based on 144 sentences randomly selected from the National Corpus of the Russian Language (RNC, ruscorpora.ru) - an online corpus of Russian texts with extensive search options. These sentences were slightly edited: the authors replaced rare infrequent words with more frequent ones and shortened the sentences when they exceeded the preset maximum length of 13 words. The stimuli sentences were subjected to the cloze task experiment. Respondents were asked to successively predict the next words for each context. An example of stimuli that were shown to the participants is presented in Table 1 with corresponding correct next words and calculated predictability scores.

@@

The Russian News Corpus \cite{runews} includes newspaper articles published in the 2000s.
The National Corpus of Russian Language \cite{natcorp} includes written texts from the middle of the 18th to the middle of the 20th century.

Some corpus statistics are presented in Table 2.

@@

All of our models (except LSTM) were trained separately on the two corpora, and on their combination. Thus we examine these models:

@@

The custom LSTM model was trained on a blinded NCRL (excluding the sentences chosen for stimuli) and the RNC. Overall, the training corpus consisted of 577 million tokens. The model was tested on 1000 sentences from the OpenCorpora project \cite{bocharov:2011} with 1,9 million tokens from newspaper articles, Russian Wikipedia, texts from blogs, fiction, non-fiction, and legal documents.

@@

\begin{enumerate}[1)]

\item        Extracts first 10 words (we have decided that that is the maximum amount of words for each context that would have meaningful probabilities as all of the words after the first ten for each model have their probabilities tends to zero) for each of 1219 contexts;

% TODO arithmetic should be TeX-ified
\item        Computes the mean probability of remaining words;

\item        Counts the difference between the probabilities of the first and last word and then the difference between mean probabilities of the previous word and the next one;

\item        Decides what amount of words vectors to use for each context based on how different the first-last word difference and the mean probability difference is – if the latter is lesser than the former, the function checks the next difference, if not – the previous amount of words would be used for semantic comparison.

\end{enumerate}

After the decision on the number of words was made, with the help of \texttt{gensim} in Python a vector for each word was extracted.
